Synthesizing Symmetric Lenses

### 2018-07-11 papers/SymmetricLenses/main.tex
pass through section 2

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,"","unk","") 
                                                    . disconnect("",company,"","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

@@

The expected number of bits required to recover data is a well-known concept in
information theory: it is the \emph{entropy} of the data~\cite{Shannon1948}.
Calculating the entropy of data requires a probability distribution over the
space of possible values for the data. Specifically, given a set $S$ and a
probability distribution over that set $P \OfType S \rightarrow \Reals$, the
entropy of $S$ with probability distribution $P$ is $-\Sigma_{s\in
  S}P(s)\Log_2P(s)$.

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}~\cite{stoch-def,stoch-rnn} (SREs), 
which are regular expressions where each operator is annotated with a
probability.  A stochastic regular expression
specifies both a set of strings and a probability
distribution over it.  From this information, we can calculate the
entropy of the SRE.

@@

At a high level, \GreedySynth looks for lenses between a given pair of
stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the highest
likelihood by performing a type-directed search. In more detail, \GreedySynth
first normalizes the stochastic regular expressions provided by \Expand into
\emph{stochastic DNF regular expressions}. These stochastic DNF regular
expressions are, intuitively, stochastic regular expressions with disjunctions
distributed over concatenations, where associativity information has been
removed. Then, \GreedySynth uses the syntax of these n-ary DNF regular
expressions to find highly normalized lenses, in a form we call \emph{simple
  symmetric n-ary DNF lenses}. These lenses include neither a composition
operator nor a type equivalence rule. These restrictions mean that each simple
symmetric n-ary DNF lens is typed by a single pair of stochastic n-ary DNF
regular expressions and \GreedySynth's search space for a given pair of regular
expressions is finite. The variations that \GreedySynth considers include
choosing between \IdentityLens{} and \Disconnect lenses and deciding how to
match multiple occurrences of the same kind of data, \EG, the two occurrences of
\lstinline{name} in the management and HR data formats (we want to map first
names to first names and last names to last names). Finally, \GreedySynth
converts this lens back into a simple symmetric lens by converting the n-ary
forms into the binary forms provided in the surface language.

>>>>

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,"","unk","") 
                                                    . disconnect("",company,"","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens, written in our language for defining simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate lenses with types, ensuring that the lens
    operates on strings that match the regular expressions. In this work, we
    \textit{synthesize} the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

@@

The expected number of bits required to recover data is the well-known information-theoretic concept 
of the \emph{entropy} of the data~\cite{Shannon1948}.
Calculating entropy requires a probability distribution over the
space of possible values for the data. Specifically, given a set $S$ and a
probability distribution $P \OfType S \rightarrow \Reals$ over $S$, the
entropy is $-\Sigma_{s\in  S}P(s)\Log_2P(s)$.

In our setting, we already have a way of expressing sets of data:
regular expressions.  What we need is a way to express probability
distributions over that data.
To that end, we adopt \emph{stochastic regular expressions}~\cite{stoch-def,stoch-rnn} (SREs), 
which are regular expressions in which each operator is annotated with a
probability.  A stochastic regular expression
specifies both a set of strings and a probability
distribution over that set, enabling us to calculate the
entropy of the SRE.

@@

At a high level, \GreedySynth looks for lenses between a given pair of
stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the highest
likelihood by performing a type-directed search. In more detail, \GreedySynth
first normalizes the stochastic regular expressions provided by \Expand into
\emph{stochastic DNF regular expressions}. These stochastic DNF regular
expressions are, intuitively, stochastic regular expressions with disjunctions
distributed over concatenations, where associativity information has been
removed. Then, \GreedySynth uses the syntax of these n-ary DNF regular
expressions to find highly normalized lenses, in a form we call \emph{simple
  symmetric n-ary DNF lenses}. These lenses include neither a composition
operator nor a type equivalence rule. These restrictions mean that each simple
symmetric n-ary DNF lens is typed by a single pair of stochastic n-ary DNF
regular expressions and \GreedySynth's search space for a given pair of regular
expressions is finite. The variations that \GreedySynth considers include
choosing between \IdentityLens{} and \Disconnect lenses and deciding how to
match multiple occurrences of the same kind of data, \EG, the two occurrences of
\lstinline{name} in the management and HR data formats (we want to map first
names to first names and last names to last names). \GreedySynth
converts yields a simple symmetric lens by converting the n-ary
forms into the binary forms provided in the surface language.
